Underdispersion in the context of community athletics

### 2024-03-28 inst/parkrun_idea.Rmd
reparametrization

f <- function(n,a){
  p <- 1:n
  p <- (p-min(p))/(max(p)-min(p))
  jj <- exp(a*p)
  jj/sum(jj)  
}

@@

alpha_try <- 0:6
beta_try <- 1:26

M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
system.time(L <- t(apply(M,1,getlike)))
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
colnames(L) <- c(letters,'alpha','beta')
head(L)

```

## Now 10 in-silico races:
```{r doinsilico,cache=TRUE}
alphatrue <- 3
betatrue <- 4
set.seed(0)
D <- replicate(30,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
```

```{r trylike,cache=TRUE}
likelihood <- apply(L[,D],1,prod)
likelihood <- matrix(likelihood,length(alpha_try),length(beta_try))
```

```{r showlike}
matplot(likelihood)
filled.contour(alpha_try,beta_try,(likelihood))
table(likelihood==0)
plot(f(26,alphatrue))
abline(v=betatrue)
```

@@

``Parkrun'' is a distributed community initiative that organises
weekly timed 5 km runs/walks in parks worldwide~\cite{hindley2020}.  A
typical event will have 200 participants.  Table~\ref{parkruntable}
shows the author has completed a total of 15 parkruns to date, and
from the first pair of numbers we see that 238 runners attended that
particular parkrun, of whom the author placed 173.

@@

We may consider the author to have an unknown generalized Bradley
Terry strength $a$: Figure~\ref{parkrunsupport} shows a support curve
for $a$ with the evaluate, $\hat{a}\simeq 0.448$ shown.  We also see
two units of support~\cite{edwards1992} shown as a dotted line
illustrating a support interval of about 0.315 - 0.567; we may be
reasonably confident that the author's true strength lies in this
range.  In particular, note that a reasonable $H_0\colon
a=\frac{1}{2}$ may not be rejected, the support for $H_0$ being only
susceptible to very minor improvement [$\simeq 0.351$] by moving to
the evaluate.

>>>>

f <- function(n,alpha){
  jj <- alpha^(1:n)
  return(jj/sum(jj))
}

@@

alpha_try <- seq(from=1,to=6,len=10)
beta_try <- 1:26

M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
system.time(L <- t(apply(M,1,getlike)))
L_integer <- L  # counts
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
colnames(L) <- c(letters,'alpha','beta')
head(L)
```

```{r makecontL}
colSums(L)
mL <- function(i){matrix(L[,i],length(alpha_try),length(beta_try))}
filled.contour(alpha_try,beta_try,mL(1))
filled.contour(alpha_try,beta_try,mL(5))
filled.contour(alpha_try,beta_try,mL(10))
filled.contour(alpha_try,beta_try,mL(15))
filled.contour(alpha_try,beta_try,mL(20))
filled.contour(alpha_try,beta_try,mL(26))
```

```{r}
plot(table(L_integer[,1:26]))
```

## Now 10 in-silico races:
```{r doinsilico,cache=FALSE}
alphatrue <- 1.3
betatrue <- 20   # 13th letter = "m"
set.seed(1)
D <- replicate(10,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
D <- 10:20
D
```

```{r trylikeinsilico,cache=FALSE}
likelihood <- apply(L[,D],1,prod) + 0.00
likelihood <- matrix(likelihood,length(alpha_try),length(beta_try))
```

```{r showlike}
round(alpha_try,2)
matplot(alpha_try,likelihood,type='b')
matplot(beta_try,t(likelihood),type='b')
matplot(alpha_try,log(likelihood)-log(max(likelihood)),type='b',ylim=c(-20,0))
matplot(beta_try,log(t(likelihood))-log(max(likelihood)),type='b',ylim=c(-20,0))
filled.contour(alpha_try,beta_try,(likelihood))
table(likelihood==0)
plot(f(26,alphatrue))
abline(v=betatrue)
contour(alpha_try,beta_try,likelihood)
abline(v=alphatrue)
abline(h=betatrue)
contour(alpha_try,beta_try,log(likelihood) - log(max(likelihood)),levels=-(0:10))
abline(v=alphatrue)
abline(h=betatrue)
```

@@

"Parkrun" is a distributed community initiative that organises
weekly timed 5 km runs/walks in parks worldwide~\cite{hindley2020}.  A
typical event will have 200 participants.  Table~\ref{parkruntable}
shows the author has completed a total of 15 parkruns to date, and
from the first pair of numbers we see that 238 runners attended that
particular parkrun, of whom the author placed 173.

@@

We may consider the author to have an unknown generalized Bradley
Terry strength $a$: Figure~\ref{parkrunsupport} shows a support curve
for $a$ with the evaluate, $\hat{a}\simeq 0.448$ shown.  We also see
two units of support~\cite{edwards1992} shown as a dotted line
illustrating a support interval of about 0.315 - 0.567; we may be
reasonably confident that the author's true BT strength lies in this
range.  In particular, note that a reasonable $H_0\colon
a=\frac{1}{2}$ may not be rejected, the support for $H_0$ being only
susceptible to very minor improvement [$\simeq 0.351$] by moving to
the evaluate.
